Taking Linear Logic Apart

### 2018-04-27 doc/tlla2018/main.tex
Minor additions

\documentclass[draft,submission,copyright,creativecommons]{eptcs}
\providecommand{\event}{Linearity and TLLA 2018}
\usepackage[british]{babel}
\usepackage[all]{foreign}
\usepackage[fleqn]{amsmath}
\usepackage{amsthm}
\allowdisplaybreaks
\usepackage[inline]{enumitem}
\usepackage{breakurl} % Not needed if you use pdflatex only.
\usepackage{underscore} % Only needed if you use pdflatex.
\usepackage{textcomp,textgreek,upgreek}
\usepackage{nameref,hyperref}
\usepackage[capitalise,nameinlink,noabbrev]{cleveref}
\hypersetup{hidelinks,final}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{example}{Example}
\input{preamble-typing}
%\usepackage[parfill]{parskip}
\usepackage{more-space}
\usepackage[sectionbib,square,numbers,sort&compress]{natbib}

@@

The term language of \cp is a variant of the \textpi-calculus.
The variables $\tm{x}$, $\tm{y}$, and $\tm{z}$ range over channel names.
The construct $\tm{\cpLink{x}{y}}$ links two
channels~\cite{sangiorgi1996,boreale1998}, forwarding messages received on
$\tm{x}$ to $\tm{y}$ and vice versa.
The construct $\tm{\cpCut{x}{P}{Q}}$ creates a new channel $\tm{x}$, and composes
two processes, $\tm{P}$ and $\tm{Q}$, which communicate on $\tm{x}$, in parallel.
Therefore, in $\tm{\cpCut{x}{P}{Q}}$ the name $\tm{x}$ is bound in both $\tm{P}$
and $\tm{Q}$.
In $\tm{\cpRecv{x}{y}{P}}$ and $\tm{\cpSend{x}{y}{P}{Q}}$, round brackets denote
input, square brackets denote output.
We use bound output~\cite{sangiorgi1996}, meaning that both input and output
bind a new name.
In $\tm{\cpRecv{x}{y}{P}}$ the new name $\tm{y}$ is bound in $\tm{P}$.
In $\tm{\cpSend{x}{y}{P}{Q}}$, the new name $\tm{y}$ is only bound in $\tm{P}$,
while $\tm{x}$ is only bound in $\tm{Q}$.
\begin{definition}[Terms]\label{def:cp-terms}  Process terms are given by the following grammar:
  \[
    \begin{array}{rllrll}
      \tm{P}, \tm{Q}, \tm{R}
           :=& \tm{\cpLink{x}{y}}       &\text{link}
      &  \mid& \tm{\cpCut{x}{P}{Q}}     &\text{parallel composition, ``cut''}
      \\ \mid& \tm{\cpSend{x}{y}{P}{Q}} &\text{output}
      &  \mid& \tm{\cpRecv{x}{y}{P}}    &\text{input}
      \\ \mid& \tm{\cpHalt{x}}          &\text{halt}
      &  \mid& \tm{\cpWait{x}{P}}       &\text{wait}
      \\ \mid& \tm{\cpInl{x}{P}}        &\text{select left choice}
      &  \mid& \tm{\cpInr{x}{P}}        &\text{select right choice}
      \\ \mid& \tm{\cpCase{x}{P}{Q}}    &\text{offer binary choice}
      &  \mid& \tm{\cpAbsurd{x}}        &\text{offer nullary choice}
    \end{array}
  \]
\end{definition}\noindent

Terms in \cp are identified up to structural congruence, which states that
links are symmetric, and parallel compositions $\tm{\cpCut{x}{P}{Q}}$ are
associative and commutative.
\begin{definition}[Structural congruence]\label{def:cp-equiv}
  The structural congruence $\equiv$ is the congruence closure over terms which
  satisfies the following additional axioms:
  \[
    \setlength{\arraycolsep}{3pt}
    \begin{array}{llcll}
      \cpEquivLinkComm
      & \tm{\cpLink{x}{y}}
      & \equiv
      & \tm{\cpLink{y}{x}}
      \\
      \cpEquivCutComm
      & \tm{\cpCut{x}{P}{Q}}
      & \equiv
      & \tm{\cpCut{x}{Q}{P}}
      \\
      \cpEquivCutAss1
      & \tm{\cpCut{x}{P}{\cpCut{y}{Q}{R}}}
      & \equiv
      & \tm{\cpCut{y}{\cpCut{x}{P}{Q}}{R}}
      & \text{if }\notFreeIn{x}{R}\text{ and }\notFreeIn{y}{P}
    \end{array}
  \]
\end{definition}\noindent

\begin{figure}
 \[
    \begin{array}{llcll}
      \cpRedAxCut1
      & \tm{\cpCut{x}{\cpLink{w}{x}}{P}}
      & \Longrightarrow
      & \tm{\cpSub{w}{x}{P}} 
      \\
      \cpRedBetaTensParr
      & \tm{\cpCut{x}{\cpSend{x}{y}{P}{Q}}{\cpRecv{x}{y}{R}}}
      & \Longrightarrow
      & \tm{\cpCut{y}{P}{\cpCut{x}{Q}{R}}}
      \\
      \cpRedBetaOneBot
      & \tm{\cpCut{x}{\cpHalt{x}}{\cpWait{x}{P}}}
      & \Longrightarrow
      & \tm{P}
      \\
      \cpRedBetaPlusWith1
      & \tm{\cpCut{x}{\cpInl{x}{P}}{\cpCase{x}{Q}{R}}}
      & \Longrightarrow
      & \tm{\cpCut{x}{P}{Q}}
      \\
      \cpRedBetaPlusWith2
      & \tm{\cpCut{x}{\cpInr{x}{P}}{\cpCase{x}{Q}{R}}}
      & \Longrightarrow
      & \tm{\cpCut{x}{P}{R}}
      \\
      \cpRedKappaTens1
      & \tm{\cpCut{x}{\cpSend{y}{z}{P}{Q}}{R}}
      & \Longrightarrow \;
      & \tm{\cpSend{y}{z}{\cpCut{x}{P}{R}}{Q}}
        \quad \text{if} \; \notFreeIn{x}{Q}
      \\
      \cpRedKappaTens2
      & \tm{\cpCut{x}{\cpSend{y}{z}{P}{Q}}{R}}
      & \Longrightarrow \;
      & \tm{\cpSend{y}{z}{P}{\cpCut{x}{Q}{R}}}
        \quad \text{if} \; \notFreeIn{x}{P}
      \\
      \cpRedKappaParr
      & \tm{\cpCut{x}{\cpRecv{y}{z}{P}}{R}}
      & \Longrightarrow \;
      & \tm{\cpRecv{y}{z}{\cpCut{x}{P}{R}}}
      \\
      \cpRedKappaBot
      & \tm{\cpCut{x}{\cpWait{y}{P}}{R}}
      & \Longrightarrow \;
      & \tm{\cpWait{y}{\cpCut{x}{P}{R}}}
      \\
      \cpRedKappaPlus1
      & \tm{\cpCut{x}{\cpInl{y}{P}}{R}}
      & \Longrightarrow \;
      & \tm{\cpInl{y}{\cpCut{x}{P}{R}}}
      \\
      \cpRedKappaPlus2
      &\tm{\cpCut{x}{\cpInr{y}{P}}{R}}
      & \Longrightarrow \;
      & \tm{\cpInr{y}{\cpCut{x}{P}{R}}}
      \\
      \cpRedKappaWith
      & \tm{\cpCut{x}{\cpCase{y}{P}{Q}}{R}}
      & \Longrightarrow \;
      & \tm{\cpCase{y}{\cpCut{x}{P}{R}}{\cpCut{x}{Q}{R}}}
      \\
      \cpRedKappaTop
      & \tm{\cpCut{x}{\cpAbsurd{y}}{R}}
      & \Longrightarrow \;
      & \tm{\cpAbsurd{y}}
    \end{array}
  \]
  \begin{center}
    \begin{prooftree*}
      \AXC{$\reducesto{P}{P^\prime}$}
      \SYM{\cpRedGammaCut}
      \UIC{$\reducesto{\cpCut{x}{P}{Q}}{\cpCut{x}{P^\prime}{Q}}$}
    \end{prooftree*}
    \begin{prooftree*}
      \AXC{$\tm{P}\equiv\tm{Q}$}
      \AXC{$\reducesto{Q}{Q^\prime}$}
      \AXC{$\tm{Q^\prime}\equiv\tm{P^\prime}$}
      \SYM{\cpRedGammaEquiv}
      \TIC{$\reducesto{P}{P^\prime}$}
    \end{prooftree*}
  \end{center}
  \caption{Classical Processes, reduction relation.}
  \label{fig:cp-reduction}
\end{figure}

Reductions relate processes with their reduced forms \eg, a reduction $\reducesto{P}{Q}$ denotes that the process $\tm{P}$ can reduce to the process $\tm{Q}$ in a single step.
\begin{definition}[Reduction]\label{def:cp-reduction}
  Reductions are described by the smallest relation $\Longrightarrow$ on process terms closed under rules in \cref{fig:cp-reduction}.
  The relation $\Longrightarrow^\star$ is the reflexive, transitive closure of 
  $\Longrightarrow$.
\end{definition}\noindent

Channels in \cp are typed using a session type system which corresponds to
classical linear logic.
A process communicating over a channel of type \ty{A \tens B} must be ready to
split into two independent processes (two processes which share no
channels), ready to communicate over a channel of type \ty{A} and \ty{B},
respectively, and process communicating over a channel of type \ty{A \parr B} is
dual to this, and therefore must be ready to start communicating concurrently
with two independent processes.
A process communicating over a channel of type \ty{A \plus B} can choose to
communicate either as \ty{A} or \ty{B}, and a process communicating over a
channel of type \ty{A \with B} must be able to communicate either as \ty{A} or
\ty{B}.
\begin{definition}[Types]\label{def:cp-types}

@@

    \begin{array}{rclrcl}
      \ty{A}, \ty{B}, \ty{C}
           :=& \ty{A \tens B} &\text{pair of independent processes}
      &  \mid& \ty{\one}      &\text{unit for} \; {\tens}
      \\ \mid& \ty{A \parr B} &\text{pair of interdependent processes}
      &  \mid& \ty{\bot}      &\text{unit for} \; {\parr}
      \\ \mid& \ty{A \plus B} &\text{internal choice}
      &  \mid& \ty{\nil}      &\text{unit for} \; {\plus}
      \\ \mid& \ty{A \with B} &\text{external choice}
      &  \mid& \ty{\top}      &\text{unit for} \; {\with}

@@

Duality plays a crucial role in both linear logic and session types.
In \cp, the two endpoints of a channel are assigned dual types.
This ensures that, for instance, whenever a process \emph{sends} across a
channel, the process on the other end of that channel is waiting to
\emph{receive}.
Each type $\ty{A}$ has a dual, written $\ty{A^\bot}$.
Duality is an involutive function.
\begin{definition}[Duality]\label{def:cp-negation}

@@

    \begin{array}{lclclclclclclcl}
               \ty{(A \tens B)^\bot} & = & \ty{A^\bot \parr B^\bot}
      &\qquad& \ty{\one^\bot}        & = & \ty{\bot}
      &\qquad& \ty{(A \parr B)^\bot} & = & \ty{A^\bot \tens B^\bot}
      &\qquad& \ty{\bot^\bot}        & = & \ty{\one}
      \\       \ty{(A \plus B)^\bot} & = & \ty{A^\bot \with B^\bot}
      &\qquad& \ty{\nil^\bot}        & = & \ty{\top}
      &\qquad& \ty{(A \with B)^\bot} & = & \ty{A^\bot \plus B^\bot}
      &\qquad& \ty{\top^\bot}        & = & \ty{\nil}

@@

Environments associate channels with types.
Names in environments must be unique, and environments $\ty{\Gamma}$ and
$\ty{\Delta}$ can only be combined as $\ty{\Gamma}, \ty{\Delta}$ if
$\ty{\fv{\Gamma}} \cap \ty{\fv{\Delta}} = \varnothing$.
\begin{definition}[Environments]\label{def:cp-environments}
    \ty{\Gamma}, \ty{\Delta}, \ty{\Theta} := \tmty{x_1}{A_1}\dots\tmty{x_n}{A_n}

@@

  \begin{center}\cpInfAx \cpInfCut \end{center}
  \vspace*{-.5\baselineskip}
  \begin{center}\cpInfTens \cpInfParr \cpInfOne \end{center}
  \vspace*{-.5\baselineskip}
  \begin{center}\cpInfBot \cpInfPlus1 \cpInfPlus2\end{center}
  \vspace*{-.5\baselineskip}
  \begin{center}\cpInfWith \cpInfNil \cpInfTop \end{center}
  \caption{Classical Processes, typing.}
  \label{fig:cp}
\noindent
Typing judgements associate processes with collections of typed channels.
\begin{definition}[Typing judgements]\label{def:cp}
  A typing judgement $\seq[{ P }]{\tmty{x_1}{A_1}\dots\tmty{x_n}{A_n}}$ denotes
  that the process \tm{P} communicates along channels $\tm{x_1}\dots\tm{x_n}$
  following protocols $\ty{A_1}\dots\ty{A_n}$.
  Typing judgements are derived using rules in \cref{fig:cp}.
\cp enjoys subject reduction, progress, and termination~\cite{wadler2012}.
\begin{lemma}[Preservation for $\equiv$]\label{lem:cp-preservation-equiv}
  If $\tm{P}\equiv\tm{Q}$, then $\seq[P]{\Gamma}$ iff $\seq[Q]{\Gamma}$.
\end{lemma}
\begin{theorem}[Preservation]\label{thm:cp-preservation}
  If $\seq[P]{\Gamma}$ and $\reducesto{P}{Q}$, then $\seq[Q]{\Gamma}$.
  By induction on the derivation of $\reducesto{P}{Q}$.
\begin{theorem}[Progress]\label{thm:cp-progress}
  If $\seq[P]{\Gamma}$, then there exists a $\tm{Q}$ such that
  $\tm{P}\Longrightarrow^\star\tm{Q}$ and $\tm{Q}$ is not a cut.
  If the last rule is \textsc{Cut}, there are four cases:
  \begin{enumerate*}[label={\alph*)}]
    if either side of the cut is an axiom, we apply \cpRedAxCut1;
    if either side of the cut is itself a cut, we recursively eliminate the cut;
    if both sides are logical rules acting on the cut formula, we apply
    one of the $\beta$-rules;
    otherwise, at least one side is a logical rule acting on a formula other
    than the cut formula, in which case we apply one of the $\kappa$-rules.
  \end{enumerate*}

@@

In this section, we introduce Hypersequent Classical Processes (\hcp), a variant
of \cp which registers parallelism in the typing jugdgements using
hypersequents, allowing us to take apart the monolithic term constructors of \cp
(\eg $\tm{\cpSend{x}{y}{P}{Q}}$) into the corresponding \textpi-calculus term
constructs.
As expected, the term language of \hcp is a variant of \cp, where the term
constructs have been taken apart into \textpi-calculus primitives.

@@

\begin{definition}[Hyper-environment]\label{def:hcp-hyper-environment}
  \(
  \ty{\mathcal{G}}, \ty{\mathcal{H}} :=
  \ty{\emptyhypercontext}
  \; \mid \;
  \ty{\Gamma_1 \hsep \dots \hsep \Gamma_n}
  \)
\end{definition}\noindent

\begin{definition}[Typing judgements]\label{def:hcp}
  A typing judgement $\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ denotes
  that the process $\tm{P}$ consists of $n$ independent, but potentially
  entangled processes, each of which communicates according to its own
  protocol $\Gamma_i$.
  Typing judgements can be constructed using the inference rules in \cref{fig:hcp}.
  Each logical rule has the side condition that $\tm{x} \not\in \ty{\mathcal{G}}$.
\end{definition}\noindent

\begin{figure}
  \begin{center}\hccpInfAx \hccpInfCycle\end{center}
  \vspace*{-0.5\baselineskip}
  \begin{center}\hccpInfMix \hccpInfHalt\end{center}
  \vspace*{-0.5\baselineskip}
  \begin{center}\hccpInfBoundTens \hccpInfParr \hccpInfOne\end{center}
  \vspace*{-0.5\baselineskip}
  \begin{center}\hccpInfBot \hccpInfPlus1 \hccpInfPlus2\end{center}
  \vspace*{0.5\baselineskip}
  \begin{center}\hccpInfWith \hccpInfNil \hccpInfTop\end{center}
  \caption{Hypersequent Classical Processes, typing.}
  \label{fig:hcp}
\end{figure}

\begin{lemma}[Preservation for $\equiv$]\label{lem:hcp-preservation-equiv}
  If $\tm{P}\equiv\tm{Q}$, then $\seq[P]{\mathcal{G}}$ iff $\seq[Q]{\mathcal{G}}$.
\end{lemma}
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
\end{proof}\noindent

\begin{theorem}[Preservation]\label{thm:hcp-preservation}
  If $\seq[P]{\mathcal{G}}$ and $\reducesto{P}{Q}$, then $\seq[Q]{\mathcal{G}}$.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\reducesto{P}{Q}$.
\end{proof}\noindent

\begin{theorem}[Progress]\label{thm:hcp-progress}
  If $\seq[P]{\mathcal{G}}$, then there exists a $\tm{Q}$ such that
  $\tm{P}\Longrightarrow^\star\tm{Q}$ and $\tm{Q}$ is not a cut or a mix.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\seq[P]{\mathcal{G}}$.
  If the last rule is a cut, there are three cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if the rule under the cut is the corresponding mix, we eliminate both as in
    \cref{thm:cp-progress};
  \item
    if the rule under the cut is an unrelated mix, we rewrite by
    \hccpEquivScopeExt{}, and recursively eliminate the cut;
  \item
    if the rule under the cut is a logical rule, we apply one of the
    $\kappa\nu$-rules.
  \end{enumerate*}
  If the last rule is mix, there are two cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if the rule under the mix is a cut or another mix, we recursively eliminate it;
  \item
    if the rule under the mix is a logical rule, we apply one of the
    $\kappa{\ppar}$-rules.
  \end{enumerate*}
\end{proof}\noindent

>>>>

\documentclass[draft,submission,copyright,creativecommons]{eptcs}
\providecommand{\event}{Linearity and TLLA 2018}
\usepackage[british]{babel}
\usepackage[all]{foreign}
\usepackage[fleqn]{amsmath}
\usepackage{amsthm}
\allowdisplaybreaks
\usepackage[inline]{enumitem}
\usepackage{breakurl} % Not needed if you use pdflatex only.
\usepackage{underscore} % Only needed if you use pdflatex.
\usepackage{textcomp,textgreek,upgreek}
\usepackage{nameref,hyperref}
\usepackage[nameinlink,noabbrev]{cleveref}
\hypersetup{hidelinks,final}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{example}{Example}
\input{preamble-typing}
\usepackage{more-space}
\usepackage[sectionbib,square,numbers,sort&compress]{natbib}
\renewcommand*{\eg}{e.g.\@\xspace}
\renewcommand*{\ie}{i.e.\@\xspace}

@@

\paragraph{Terms}
The term language of \cp is a variant of the \textpi-calculus.
The variables $\tm{x}$, $\tm{y}$, and $\tm{z}$ range over channel names.
The construct $\tm{\cpLink{x}{y}}$ links two
channels~\cite{sangiorgi1996,boreale1998}, forwarding messages received on
$\tm{x}$ to $\tm{y}$ and vice versa.
The construct $\tm{\cpCut{x}{P}{Q}}$ creates a new channel $\tm{x}$, and composes
two processes, $\tm{P}$ and $\tm{Q}$, which communicate on $\tm{x}$, in parallel.
Therefore, in $\tm{\cpCut{x}{P}{Q}}$ the name $\tm{x}$ is bound in both $\tm{P}$
and $\tm{Q}$.
In $\tm{\cpRecv{x}{y}{P}}$ and $\tm{\cpSend{x}{y}{P}{Q}}$, round brackets denote
input, square brackets denote output.
We use bound output~\cite{sangiorgi1996}, meaning that both input and output
bind a new name.
In $\tm{\cpRecv{x}{y}{P}}$ the new name $\tm{y}$ is bound in $\tm{P}$.
In $\tm{\cpSend{x}{y}{P}{Q}}$, the new name $\tm{y}$ is only bound in $\tm{P}$,
while $\tm{x}$ is only bound in $\tm{Q}$.
\begin{definition}[Terms]\label{def:cp-terms}
  Process terms are given by the following grammar:

@@

      &  \mid& \tm{\cpCut{x}{P}{Q}}     &\text{parallel composition, ``cut''}
      \\ \mid& \tm{\cpSend{x}{y}{P}{Q}} &\text{output}
      &  \mid& \tm{\cpRecv{x}{y}{P}}    &\text{input}
      \\ \mid& \tm{\cpHalt{x}}          &\text{halt}
      &  \mid& \tm{\cpWait{x}{P}}       &\text{wait}

@@

Terms in \cp are identified up to structural congruence, which states that
links are symmetric, and parallel compositions $\tm{\cpCut{x}{P}{Q}}$ are
associative and commutative.
\begin{definition}[Structural congruence]\label{def:cp-equiv}

@@

      \cpEquivCutComm
      & \tm{\cpCut{x}{Q}{P}}
      \cpEquivCutAss1
      & \tm{\cpCut{x}{P}{\cpCut{y}{Q}{R}}}
      & \tm{\cpCut{y}{\cpCut{x}{P}{Q}}{R}}
      & \text{if }\notFreeIn{x}{R}\text{ and }\notFreeIn{y}{P}

@@

      \cpRedAxCut1
      \cpRedBetaTensParr
      & \tm{\cpCut{x}{\cpSend{x}{y}{P}{Q}}{\cpRecv{x}{y}{R}}}
      & \tm{\cpCut{y}{P}{\cpCut{x}{Q}{R}}}
      \cpRedBetaOneBot
      & \tm{\cpCut{x}{\cpHalt{x}}{\cpWait{x}{P}}}
      \cpRedBetaPlusWith1
      \cpRedBetaPlusWith2
      \cpRedKappaTens1
      & \tm{\cpCut{x}{\cpSend{y}{z}{P}{Q}}{R}}
      & \Longrightarrow \;
      & \tm{\cpSend{y}{z}{\cpCut{x}{P}{R}}{Q}}
        \quad \text{if} \; \notFreeIn{x}{Q}
      \cpRedKappaTens2
      & \tm{\cpCut{x}{\cpSend{y}{z}{P}{Q}}{R}}
      & \Longrightarrow \;
      & \tm{\cpSend{y}{z}{P}{\cpCut{x}{Q}{R}}}
        \quad \text{if} \; \notFreeIn{x}{P}
      \cpRedKappaParr
      & \tm{\cpCut{x}{\cpRecv{y}{z}{P}}{R}}
      & \Longrightarrow \;
      & \tm{\cpRecv{y}{z}{\cpCut{x}{P}{R}}}
      \cpRedKappaBot
      & \tm{\cpCut{x}{\cpWait{y}{P}}{R}}
      & \Longrightarrow \;
      & \tm{\cpWait{y}{\cpCut{x}{P}{R}}}
      \cpRedKappaPlus1
      & \tm{\cpCut{x}{\cpInl{y}{P}}{R}}
      & \Longrightarrow \;
      & \tm{\cpInl{y}{\cpCut{x}{P}{R}}}
      \cpRedKappaPlus2
      &\tm{\cpCut{x}{\cpInr{y}{P}}{R}}
      & \Longrightarrow \;
      & \tm{\cpInr{y}{\cpCut{x}{P}{R}}}
      \cpRedKappaWith
      & \tm{\cpCut{x}{\cpCase{y}{P}{Q}}{R}}
      & \Longrightarrow \;
      & \tm{\cpCase{y}{\cpCut{x}{P}{R}}{\cpCut{x}{Q}{R}}}
      \cpRedKappaTop
      & \tm{\cpCut{x}{\cpAbsurd{y}}{R}}
      & \Longrightarrow \;
      \SYM{\cpRedGammaCut}
      \UIC{$\reducesto{\cpCut{x}{P}{Q}}{\cpCut{x}{P^\prime}{Q}}$}
      \SYM{\cpRedGammaEquiv}
  \caption{Classical Processes, reduction relation.}
  \label{fig:cp-reduction}
Reductions relate processes with their reduced forms \eg, a reduction
$\reducesto{P}{Q}$ denotes that the process $\tm{P}$ can reduce to the process
$\tm{Q}$ in a single step.
\begin{definition}[Reduction]\label{def:cp-reduction}
  terms closed under rules in \cref{fig:cp-reduction}.
  The relation $\Longrightarrow^\star$ is the reflexive, transitive closure of 
  $\Longrightarrow$.

@@

\paragraph{Typing}
Channels in \cp are typed using a session type system which corresponds to
classical linear logic.
A process communicating over a channel of type \ty{A \tens B} must be ready to
split into two independent processes (two processes which share no
channels), ready to communicate over a channel of type \ty{A} and \ty{B},
respectively, and process communicating over a channel of type \ty{A \parr B} is
dual to this, and therefore must be ready to start communicating concurrently
with two independent processes.
A process communicating over a channel of type \ty{A \plus B} can choose to
communicate either as \ty{A} or \ty{B}, and a process communicating over a
channel of type \ty{A \with B} must be able to communicate either as \ty{A} or
\ty{B}.
\begin{definition}[Types]\label{def:cp-types}
  \[
    \begin{array}{rclrcl}
      \ty{A}, \ty{B}, \ty{C}
           :=& \ty{A \tens B} &\text{pair of independent processes}
      &  \mid& \ty{\one}      &\text{unit for} \; {\tens}
      \\ \mid& \ty{A \parr B} &\text{pair of interdependent processes}
      &  \mid& \ty{\bot}      &\text{unit for} \; {\parr}
      \\ \mid& \ty{A \plus B} &\text{internal choice}
      &  \mid& \ty{\nil}      &\text{unit for} \; {\plus}
      \\ \mid& \ty{A \with B} &\text{external choice}
      &  \mid& \ty{\top}      &\text{unit for} \; {\with}
    \end{array}
  \]
\end{definition}\noindent
Duality plays a crucial role in both linear logic and session types.
In \cp, the two endpoints of a channel are assigned dual types.
This ensures that, for instance, whenever a process \emph{sends} across a
channel, the process on the other end of that channel is waiting to
\emph{receive}.
Each type $\ty{A}$ has a dual, written $\ty{A^\bot}$.
Duality is an involutive function.
\begin{definition}[Duality]\label{def:cp-negation}
  \[
    \setlength{\arraycolsep}{3pt}
    \begin{array}{lclclclclclclcl}
               \ty{(A \tens B)^\bot} & = & \ty{A^\bot \parr B^\bot}
      &\qquad& \ty{\one^\bot}        & = & \ty{\bot}
      &\qquad& \ty{(A \parr B)^\bot} & = & \ty{A^\bot \tens B^\bot}
      &\qquad& \ty{\bot^\bot}        & = & \ty{\one}
      \\       \ty{(A \plus B)^\bot} & = & \ty{A^\bot \with B^\bot}
      &\qquad& \ty{\nil^\bot}        & = & \ty{\top}
      &\qquad& \ty{(A \with B)^\bot} & = & \ty{A^\bot \plus B^\bot}
      &\qquad& \ty{\top^\bot}        & = & \ty{\nil}
    \end{array}
  \]
\end{definition}\noindent
Environments associate channels with types.
Names in environments must be unique, and environments $\ty{\Gamma}$ and
$\ty{\Delta}$ can only be combined as $\ty{\Gamma}, \ty{\Delta}$ if
$\ty{\fv{\Gamma}} \cap \ty{\fv{\Delta}} = \varnothing$.
\begin{definition}[Environments]\label{def:cp-environments}
  \(
    \ty{\Gamma}, \ty{\Delta}, \ty{\Theta} := \tmty{x_1}{A_1}\dots\tmty{x_n}{A_n}
  \) 
\end{definition}\noindent
\begin{figure}
  \begin{center}\cpInfAx \cpInfCut \end{center}
  \vspace*{-.5\baselineskip}
  \begin{center}\cpInfTens \cpInfParr \cpInfOne \end{center}
  \vspace*{-.5\baselineskip}
  \begin{center}\cpInfBot \cpInfPlus1 \cpInfPlus2\end{center}
  \vspace*{-.5\baselineskip}
  \begin{center}\cpInfWith \cpInfNil \cpInfTop \end{center}
  \caption{Classical Processes, typing.}
  \label{fig:cp}
\end{figure}\noindent
Typing judgements associate processes with collections of typed channels.
\begin{definition}[Typing judgements]\label{def:cp}
  A typing judgement $\seq[{ P }]{\tmty{x_1}{A_1}\dots\tmty{x_n}{A_n}}$ denotes
  that the process \tm{P} communicates along channels $\tm{x_1}\dots\tm{x_n}$
  following protocols $\ty{A_1}\dots\ty{A_n}$.
  Typing judgements are derived using rules in \cref{fig:cp}.
\end{definition}\noindent

\paragraph{Metatheory}
\cp enjoys subject reduction, progress, and termination~\cite{wadler2012}.
\begin{lemma}[Preservation for $\equiv$]\label{lem:cp-preservation-equiv}
  If $\tm{P}\equiv\tm{Q}$, then $\seq[P]{\Gamma}$ iff $\seq[Q]{\Gamma}$.
\end{lemma}
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
\end{proof}\noindent
\begin{theorem}[Preservation]\label{thm:cp-preservation}
  If $\seq[P]{\Gamma}$ and $\reducesto{P}{Q}$, then $\seq[Q]{\Gamma}$.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\reducesto{P}{Q}$.
\end{proof}\noindent
\begin{theorem}[Progress]\label{thm:cp-progress}
  If $\seq[P]{\Gamma}$, then there exists a $\tm{Q}$ such that
  $\tm{P}\Longrightarrow^\star\tm{Q}$ and $\tm{Q}$ is not a cut.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\seq[P]{\Gamma}$.
  If the last rule is \textsc{Cut}, there are four cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if either side of the cut is an axiom, we apply \cpRedAxCut1;
  \item
    if either side of the cut is itself a cut, we recursively eliminate the cut;
  \item
    if both sides are logical rules acting on the cut formula, we apply
    one of the $\beta$-rules;
  \item
    otherwise, at least one side is a logical rule acting on a formula other
    than the cut formula, in which case we apply one of the $\kappa$-rules.
  \end{enumerate*}
\end{proof}\noindent
\begin{theorem}[Termination]\label{thm:cp-termination}
  If $\seq[P]{\Gamma}$, then there are no infinite $\Longrightarrow$-reduction
  sequences.
\end{theorem}
\begin{proof}
  Every reduction reduces a single cut to zero, one or two cuts.
  However, each of these cuts is smaller, in the sense that the type of the
  channel on which the communication takes place is smaller.
  Each reduction either eliminates a connective.
  Furthermore, each instance of the structural congruence preserves the size of
  the cut.
  Therefore, there cannot be an infinite $\Longrightarrow$-reduction sequence.
\end{proof}

@@

In this section, we introduce Hypersequent Classical Processes (\hcp), a variant
of \cp which registers parallelism in the typing jugdgements using
hypersequents, allowing us to take apart the monolithic term constructors of \cp
(\eg $\tm{\cpSend{x}{y}{P}{Q}}$) into the corresponding \textpi-calculus term
constructs.

\paragraph{Terms}
The term language of \hcp is a variant of \cp where the term constructs have
been taken apart into \textpi-calculus primitives.

@@

\paragraph{Typing}
We use the same definitions for types and environments for \hcp as we used for \cp.
However, on top of this, we add the definition of a hyper-environment.
Hyper-environments are either empty, or consist of a series of typing
environments, separated by vertical bars.
A hyper-environment $\ty{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ can be used to
type a series of $n$ entangled, but independent processes.
\begin{definition}[Hyper-environment]\label{def:hcp-hyper-environment}
  \(
  \ty{\mathcal{G}}, \ty{\mathcal{H}} :=
  \ty{\emptyhypercontext}
  \; \mid \;
  \ty{\Gamma_1 \hsep \dots \hsep \Gamma_n}
  \)
\end{definition}\noindent
Typing judgements in \hcp associate processes with hyper-environments.
\textsc{H-Mix} composes two processes in parallel, but remembers that they are
independent on in the sequent.
\textsc{H-Cut} and $(\tens)$ take as their premise a process which consists of
at least two unconnected processes, and connects them, eliminating the vertical
bar.
Note that each logical rule has the side condition that $\tm{x} \not\in
\mathcal{G}$, which can be read as ``you cannot act on one end-point of $x$ if
you are also holding its other end-point''.
This prevents self-locking processes, \eg
$\tm{\piBoundSend{x}{}{\piRecv{x}{}{\piHalt}}}$.
\begin{definition}[Typing judgements]\label{def:hcp}
  A typing judgement $\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ denotes
  that the process $\tm{P}$ consists of $n$ independent, but potentially
  entangled processes, each of which communicates according to its own
  protocol $\Gamma_i$.
  Typing judgements can be constructed using the inference rules in \cref{fig:hcp}.
  Furthermore, each logical rule has the side condition that $\tm{x} \not\in
  \ty{\mathcal{G}}$.
\end{definition}\noindent
\begin{figure}
  \begin{center}\hccpInfAx \hccpInfCycle\end{center}
  \vspace*{-0.5\baselineskip}
  \begin{center}\hccpInfMix \hccpInfHalt\end{center}
  \vspace*{-0.5\baselineskip}
  \begin{center}\hccpInfBoundTens \hccpInfParr \hccpInfOne\end{center}
  \vspace*{-0.5\baselineskip}
  \begin{center}\hccpInfBot \hccpInfPlus1 \hccpInfPlus2\end{center}
  \vspace*{0.5\baselineskip}
  \begin{center}\hccpInfWith \hccpInfNil \hccpInfTop\end{center}
  \caption{Hypersequent Classical Processes, typing.}
  \label{fig:hcp}
\end{figure}
